\mythesistitleNOTSYNPOSIS \tnote[fyi]{Synopsis title is ``Generalization of Consecutive Ones Property of Binary Matrices''}

### 2012-02-20 src/030chap3.tex
- incorporated new changes to Characterization algorithm. no more exit    statements in algo 1 and 2. however it is not identitcal to n's    version. kept some of the old lemmas because it seems to elucidate    (to me) the thesis. - partial editing done of general tree section (prime matrix theory)    -- TBD - not sure about status of special tree = path section -- TBD
- changed notation for induced path labeling
- changed filter 1 and 2 names.
- poset definition

Section~\ref{sec:myresearchintro} recalls the idea of tree path
labeling. The necessary preliminaries with definitions etc. are
presented in Section~\ref{ch:prelims}. Section~\ref{sec:feasible}
documents the characterization of a feasible path labeling from any
target trees. Section~\ref{sec:spltargettree} describes two special
cases where the target tree is of a particular family of trees. The
first one is shown to be equivalent to COP testing in
Section~\ref{sec:icpplicpia}. Section~\ref{sec:ksubdivstar} discusses
the second special case and presents a polynomial time algorithm to
find the tree path labeling of a given set system from a given
\kstar. Section~\ref{sec:norestraint} discusses the plain vanilla
version where the target tree has no restrictions and the algorithm to
find a feasible TPL, if any, in this case.

@@

\begin{definition}[In-tree]
An {\em in-tree} is a directed rooted tree in which all edges are
directed toward to the root.  
\end{definition}
In-trees are used in Section~\ref{sec:decompos} which describes our
solution to \CFTPL problem.

@@

In the remaining part of this section we show that $(\cF, \cl)$ is
feasible if and only if it is an ICPPL%  and
% Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty path
% hypergraph isomorphism
% function
.  Algorithm~\ref{al:icppl-find-isomorph} takes $(\cF, \cl)$ as input
and computes the feasibility bijection if it is feasible or else
reports failure. This algorithm recursively does two levels of
filtering of $(\cF, \cl)$ to make it simpler in terms of set
intersections while retaining a non-empty subset of isomorphisms, if
any, between $\cF$ and $\cF^\cl$. These two filters are {\tt
  filter\_1} in Algorithm~\ref{al:filterone} and {\tt filter\_2} in
Algorithm~\ref{al:filtertwo}.
% One direction of this claim isclear: that if a path labeling is
% feasible, then all intersection cardinalities are preserved,
% i.e. the path labeling is an ICPPL. Algorithm~\ref{al:filterone}
% \annote{has no premature exit condition hence any input will go
%   through it}{Prove that the filtered sets has ICPPL \iff input PL
%   has ICPPL?}. Algorithm~\ref{al:filtertwo} has an exit condition at
% line~\ref{xempty}. It can be easily verified that $X$ cannot be
% empty if $\cl$ is a feasible path labeling. The reason is that a
% feasible path labeling has an associated bijection between
% $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that the
% sets map to paths, ``preserving'' the path labeling.  The rest of
% the section is devoted to constructively proving that it is
% sufficient for a path labeling to be an ICPPL and pass the two
% filtering algorithms.  To describe in brief, the constructive
% approaches refine an ICPPL iteratively, such that at the end of each
% iteration we have a ``filtered'' path labeling, and finally we have
% a path labeling that defines a family of bijections from $supp(\cF)$
% to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.

First, we present Algorithm~\ref{al:filterone} which describes {\tt
  filter\_1}, and prove its correctness.  This algorithm refines the
path labeling by processing pairs of paths in $\cF^\cl$ that share a
leaf until no two paths in the new path labeling share any leaf.

\begin{algorithm}[h]
  \caption{Refine ICPPL {\tt filter\_1($\cF, \cl, T$)}}
  \label{al:filterone} % {perms}
  \begin{algorithmic}[\lndisplay]
    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$\\
    \STATE $j \assign 1$\\
    \WHILE {there is $S_1, S_2 \in \cF_{j-1}$ such that
      $\cl_{j-1}(S_1)$ and $\cl_{j-1}(S_2)$ have a common leaf in
      $T$}\label{shareleaf} \STATE $\cF_j \assign (\cF_{j-1} \setminus
    \{S_1, S_2\}) \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2
    \setminus S_1 \}$ \label{setbreak} \COMMENT {Remove $S_1$, $S_2$
      and add the ``filtered'' sets} \STATE {\bf for} every $S \in
    \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
    S_2$ {\bf do} $\cl_j(S) \assign \cl_{j-1}(S)$ {\bf end for}\\

@@

The following lemma is useful in subsequent arguments.
\begin{lemma}
  \label{lem:setminuscard}
  If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
  (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
  \cl(S_3))|$.
\end{lemma}
\begin{proof}\thesisspacing%[Proof of Lemma~\ref{lem:setminuscard}]
  Let $P_i = \cl(S_i)$, for all $1 \le i \le 3$.  $|S_1 \cap (S_2
  \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| = |S_1 \cap S_2| -
  |S_1 \cap S_2 \cap S_3|$. Due to properties (ii) and (iii) of ICPPL,
  $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1 \cap P_2| - |P_1
  \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3| = |P_1 \cap (P_2
  \setminus P_3)|$. %\qed
\end{proof}

Now we will prove the effectiveness of Algorithm~\ref{al:filterone}.
Lemma~\ref{lem:feasible} proves that {\tt filter\_1} does not destroy
the feasibility of the input, if it is indeed feasible. Following
this, Lemma~\ref{lem:invar1} and Lemma~\ref{lem:noexit1} show that if
the input is an ICPPL the result of {\tt filter\_1} is also an ICPPL
and they both have the same set of feasibility hypergraph
isomorphisms.

\begin{lemma}
  \label{lem:feasible}
  In Algorithm~\ref{al:filterone}, if input $(\cF, \cl)$ is a feasible path
  labeling then at the end of $j$th iteration of the {\bf while}
  loop, $j \ge 0$, $(\cF_j, \cl_j)$ is a feasible path labeling.
\end{lemma}
\begin{proof}\thesisspacing%[Proof of Lemma~\ref{lem:feasible}]
  We will prove this by mathematical induction on the number of
  iterations. The base case $(\cF_0, \cl_0)$ is feasible since it is
  the input itself which is given to be feasible. Assume the lemma is
  true till $j-1$th iteration \ie every hypergraph isomorphism $\phi:
  supp\left(\cF_{j-1}\right) \rightarrow V\left(T \right)$ that
  defines $(\cF, \cl)$'s feasibility, is such that the induced path
  labeling on $\cF_{j-1}$, say denoted by $\cp_{\phi[{\cF_{j-1}}]}$,
  is equal to $\cl_{j-1}$.
% Note that any hypergraph isomorphism of
% $\cF$ (say, one that defines the feasibility of $(\cF,\cl)$) can be
% ``projected'' to $\cF_{j-1}$ since their supports are equal. If $\phi$
% is a hypergraph isomorphism of to $(\cF, \cl)$, then 
  We will prove that $\phi$ is also the bijection that makes $(\cF_j,
  \cl_j)$ feasible. Note that $supp(\cF_{j-1}) = supp(\cF_{j})$ since
  the new sets in $\cF_j$ are created from basic set operations to the
  sets in $\cF_{j-1}$ adding or removing no elements. For the same
  reason and $\phi$ being a bijection, it is clear that when applying
  the $\phi$-induced path labeling on $\cF_j$, $
  \cp_{\phi[{\cF_{j}}]}(S_1 \setminus S_2) =
  \cp_{\phi[{\cF_{j-1}}]}(S_1) \setminus
  \cp_{\phi[{\cF_{j-1}}]}(S_2)$. Now observe that $ \cl_j(S_1
  \setminus S_2) = \cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2) =
  \cp_{\phi[{\cF_{j-1}}]}(S_1) \setminus
  \cp_{\phi[{\cF_{j-1}}]}(S_2)$. Thus the induced path labeling
  $\cp_{\phi[{\cF_{j}}]} = \cl_{j}$.  %\qed
\end{proof}

@@

    \tnote[dire]{Due to the condition in line~\ref{ln:3waycheck}, this invariant is
    ensured at the end of every iteration.}
    % \begin{enumerate} [{Case 3.}1:]
    % \item {\em $R$, $R'$ and $R''$ are not new sets.} Trivially true
    %   by induction hypothesis.
    % \item {\em Only one, say $R$, is a new set.}  If $R = S_1 \cap
    %   S_2$, from Lemma~\ref{lem:fourpaths} and invariant III
    %   hypothesis, this case is proven. Similarly if $R$ is any of
    %   the other new sets, the case is proven by also using Lemma
    %   ~\ref{lem:setminuscard}.
    % \item {\em At least two of $R, R', R''$ are new sets.}  The new
    %   sets are disjoint hence this case is vacuously true.
    % \end{enumerate}
  \end{enumerate} %\qed
  % \vspace{-6mm}
\end{proof}

\begin{lemma}
  \label{lem:noexit1}
  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{al:filterone} is
  feasible, then the set of hypergraph isomorphism functions that
  defines $(\cF, \cl)$'s feasibility is the same as the set that
  defines $(\cF_j, \cl_j)$'s feasibility, if any.%   Secondly, for any
%   iteration $j > 0$ of the {\em \bf while} loop, the {\em \bf exit}
%   statement in line~\ref{ln:exit1} will not execute.
\end{lemma}
\begin{proof}\thesisspacing
  Since $(\cF,\cl)$ is feasible, by Lemma~\ref{lem:feasible}
  $(\cF_j,\cl_j)$ for every iteration $j > 0$ is
  feasible.  % Therefore,
  % every hypergraph isomorphism $\phi: supp(\cF) \rightarrow V(T)$
  % that induces $\cl$ on $\cF$ also induces $\cl_{j-1}$ and $\cl_{j}$
  % on $\cF_{j-1}$ and $\cF_{j}$ respectively, i.e.,
  % $\cl_{\phi[\cF_{j-1}]} = \cl_{j-1}$ and $\cl_{\phi[\cF_j]} =
  % \cl_j$. Thus it can be seen that for all $x \in supp(\cF)$, for
  % all $v \in V(T)$ the following hold true.
  Also, every hypergraph isomorphism $\phi: supp(\cF) \rightarrow
  V(T)$ that induces $\cl$ on $\cF$ also induces $\cl_{j}$ on
  $\cF_{j}$, i.e., $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
  that for all $x \in supp(\cF)$, for all $v \in V(T)$, if $(x,v) \in
  \phi$ then $v \in \cl_{j}(S)$ for all $S \in \cF_{j}$ such that $x
  \in S$.
  % the following hold true.
  % \begin{enumerate}[i. ]
  % \item If $(x,v) \in \phi$ then $v \in \cl_{j-1}(S)$ for all $S \in
  %   \cF_{j-1}$ such that $x \in S$.
  % \item If $(x,v) \in \phi$ then $v \in \cl_{j}(S)$ for all $S \in
  %   \cF_{j}$ such that $x \in S$
  % \end{enumerate}
  In other words, {\tt filter\_1} outputs a filtered path labeling that
  ``preserves''
  hypergraph isomorphisms of the original path labeling.
%   Secondly, line~\ref{ln:exit1} will execute \iff the exit condition in
%   line~\ref{ln:3waycheck}, i.e. failure of three way intersection
%   preservation, becomes true in any iteration of the {\em \bf while}
%   loop.  Due to Lemma~\ref{lem:invar1} Invariant IV, the exit
%   condition does not occur if the input is a feasible ICPPL.%\qed
  % such that $\phi(x) = v$ where $v$ is the leaf considered in the
  % first iterations of while. Clearly, $\phi$ is a renaming of
  % vertices in hypergraph $\cF$ to those in hypergraph
  % $\cF^\cl$. Thus the following facts can be observed in every
  % iteration of the loop.
  % \begin{enumerate}[\hspace{2mm}i. ] \vspace{\topshrink}
  % \item all intersection cardinalities are preserved in this path
  %   labeling \vspace{\topshrink}
  % \item element $x$ is exclusive in a hyperedge in $\cF$ since $v$
  %   is exclusive in a hyperedge in $\cF^\cl$.
  % \end{enumerate}
  % Thus the exit condition is never rendered true after $x$ and $v$
  % are removed from their respective hyperedges. %\qed

@@

As a result of {\tt filter\_1} each leaf $v$ in $T$ is such that there
is exactly one set in $\cF$ with $v$ as a vertex in the path assigned
to it.  {\tt filter\_2}, in Algorithm~\ref{al:filtertwo}, identifies
elements in $supp(\cF)$ whose images are leaves in a hypergraph
isomorphism if one exists.  Let $S \in \cF$ be such that $\cl(S)$ is a
path with leaf and $v \in V(T)$ is the unique leaf incident on it.  We
define a new path labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) =
\set{v}$ where $x$ an arbitrary element from $S \setminus \bigcup_{\hS
  \ne S} \hS$. In other words, $x$ is an element present in no other
set in $\cF$ except $S$. This is intuitive since $v$ is present in no
other path image under $\cl$ other than $\cl(S)$.  The element $x$ and
leaf $v$ are then removed from the set $S$ and path $\cl(S)$
respectively. After doing this for all leaves in $T$, all path images
in the new path labeling $\cl_{new}$ except leaf labels (a path that
has only a leaf is called the {\em leaf label} for the corresponding
single element hyperedge or set) are paths from a new pruned tree $T_0
= T \setminus \{v \mid v \text{ is a leaf in }
T\}$. Algorithm~\ref{al:filtertwo} is now presented with details.

@@

% \begin{lemma}
%   \label{lem:setminuscard}
%   If $\cA$ is an ICPPA, and $(S_1, P_1),(S_2, P_2),(S_3, P_3) \in
%   \cA$, then $|S_1 \cap (S_2 \setminus S_3)| = |P_1 \cap (P_2
%   \setminus P_3)|$.
% \end{lemma}
% \begin{proof}\thesisspacing
%   $|S_1 \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| =
%   |S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to conditions (ii)
%   and (iii) of ICPPA, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1
%   \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3|
%   = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is proven. %\qed
% \end{proof}

>>>>

Section~\ref{sec:myresearchintro} recalls the idea of tree path
labeling. The necessary preliminaries with definitions etc. are
presented in Section~\ref{ch:prelims}. Section~\ref{sec:feasible}
documents a characterization for feasible path labelings. In
Section~\ref{sec:wallacesol} we demonstrate the algorithm described in
Section~\ref{sec:feasible} by working out the solution to an example.
Section~\ref{sec:spltargettree} describes two special cases where the
target tree is of a particular family of trees. The first one is shown
to be equivalent to COP testing in
Section~\ref{sec:icpplicpia}. Section~\ref{sec:ksubdivstar} discusses
the second special case and presents a polynomial time algorithm to
find the tree path labeling of a given set system from a given
\kstar. Section~\ref{sec:norestraint} discusses the plain vanilla
version where the target tree has no restrictions and the algorithm to
find a feasible TPL, if any, in this case.  Finally, in Section
\ref{sec:tplcomplexity} we discuss the complexity of the algorithms
seen thus far and we conclude by showing that \cop testing, which is a
special case of \TPL, is in \logspace.

@@

\begin{definition}[In-tree]
An {\em in-tree} is a directed rooted tree in which all edges are
directed toward to the root.  
\end{definition}
In-trees are used in Section~\ref{sec:norestraint} which describes our
(albeit non-polynomial time) solution to \CFTPL problem.

@@

In the remaining part of this section we show that $(\cF, \cl)$ is
feasible if and only if it is an ICPPL.
Algorithm~\ref{al:icppl-find-isomorph} takes $(\cF, \cl)$ as input and
computes the feasibility bijection if it is feasible or else reports
failure. This algorithm recursively does two levels of filtering of
$(\cF, \cl)$ to make it simpler in terms of set intersections while
retaining a non-empty subset of isomorphisms, if any, between $\cF$
and $\cF^\cl$. These two filters are \filteri in
Algorithm~\ref{al:filterone} and \filterii in
Algorithm~\ref{al:filtertwo}.
% One direction of this claim isclear: that if a path labeling is
% feasible, then all intersection cardinalities are preserved,
% i.e. the path labeling is an ICPPL. Algorithm~\ref{al:filterone}
% \annote{has no premature exit condition hence any input will go
%   through it}{Prove that the filtered sets has ICPPL \iff input PL
%   has ICPPL?}. Algorithm~\ref{al:filtertwo} has an exit condition at
% line~\ref{xempty}. It can be easily verified that $X$ cannot be
% empty if $\cl$ is a feasible path labeling. The reason is that a
% feasible path labeling has an associated bijection between
% $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that the
% sets map to paths, ``preserving'' the path labeling.  The rest of
% the section is devoted to constructively proving that it is
% sufficient for a path labeling to be an ICPPL and pass the two
% filtering algorithms.  To describe in brief, the constructive
% approaches refine an ICPPL iteratively, such that at the end of each
% iteration we have a ``filtered'' path labeling, and finally we have
% a path labeling that defines a family of bijections from $supp(\cF)$
% to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.

First, we present Algorithm~\ref{al:filterone} which describes \filteri, and prove its correctness.  This algorithm refines the
path labeling by processing pairs of paths in $\cF^\cl$ that share a
leaf until no two paths in the new path labeling share any leaf.

\begin{algorithm}[h]
  \caption{Refine ICPPL {\tt \filteri($\cF, \cl, T$)}}
  \label{al:filterone} % {perms}
  \begin{algorithmic}[\lndisplay]
    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$\\
    \STATE $j \assign 1$\\
    \WHILE {there is $S_1, S_2 \in \cF_{j-1}$ such that
      $\cl_{j-1}(S_1)$ and $\cl_{j-1}(S_2)$ have a common leaf in
      $T$}\label{shareleaf} \STATE $\cF_j \assign (\cF_{j-1} \setminus
    \{S_1, S_2\}) \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2
    \setminus S_1 \}$ \label{setbreak} \COMMENT {Remove $S_1$, $S_2$
      and add the ``filtered'' sets} \STATE {\bf for} every $S \in
    \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
    S_2$ {\bf do} $\cl_j(S) \assign \cl_{j-1}(S)$ {\bf end for}\\

@@

In order to prove the correctness of Algorithm~\ref{al:filterone}, we
present a few lemma which are useful in subsequent arguments.

The first two lemma below are from results in \cite{nsnrs09}.  They
have been reworded to use the terminology adopted in this
thesis. Lemma~\ref{lem:intersection-cardinality} shows that the
preservation of pairwise intersection cardinality in an interval
assignment is sufficient to preserve three way intersection
cardinality. Lemma~\ref{lem:icpia} shows that if a path labeling (i)
has target tree as a path \ie it is an interval labeling, and (ii) it
preserves pairwise intersection cardinalities \ie it is an ICPIA,
then it is a feasible path labeling.

\begin{lemma}[{\cite[Lem.~1]{nsnrs09}}]
  \label{lem:intersection-cardinality} 
  Let $P$ be a path, $S_1, S_2, S_3$ be 3 sets, and $T_1, T_2, T_3$ be
  paths from $P$, such that $|S_i \cap S_j |=|T_i \cap T_j|, 1 \leq
  i,j \leq 3$.  Then, $|S_1 \cap S_2 \cap S_3| = |T_1 \cap T_2 \cap
  T_3|$.
\end{lemma}
\begin{lemma}[{\cite[Th.~2]{nsnrs09}}]
  \label{lem:icpia} 
  A path labeling $(\cF, \cl)$ with target tree $T$ as a path is
  feasible iff it is an ICPIA.
\end{lemma}

Next, we see Lemma~\ref{lem:setminuscard} which is a direct
consequence of Definition~\ref{def:icppl} and is analogous to
\cite[Cor.~1]{nsnrs09}.

\begin{lemma}
  \label{lem:setminuscard}
  If $\cl$ is an \ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
  (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
  \cl(S_3))|$.
\end{lemma}
\begin{proof}\thesisspacing%[Proof of Lemma~\ref{lem:setminuscard}]
  Let $P_i = \cl(S_i)$, for all $1 \le i \le 3$.  $|S_1 \cap (S_2
  \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| = |S_1 \cap S_2| -
  |S_1 \cap S_2 \cap S_3|$. Due to properties (ii) and (iii) of ICPPL,
  $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1 \cap P_2| - |P_1
  \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3| = |P_1 \cap (P_2
  \setminus P_3)|$. %\qed
\end{proof}

For interval labeling, \cite{nsnrs09} shows that if it is an ICPIA \ie
pairwise intersection cardinalities are preserved, then $d$-way
intersection cardinalities are preseved for $2 < d < n$. Thus, in a
way, pairwise intersections ``capture'' feasibility of an interval
labeling. In the same vein, the following lemma indicates that
three-way intersection cardinality ``captures'' $d$-way intersection
cardinalities for $3 < d < n$ in a tree path labeling.

\begin{lemma}
  \label{lem:4-way}
  Let $(\cF, \cl)$ be an \ICPPL and $\cl(S_i)=P_i$, $S_i \in \cF$, $1
  \leq i \leq 4$.  Then, $|\cap_{i=1}^4 S_i| = |\cap_{i=1}^4 P_i|$.
\end{lemma}
\begin{proof}
  Consider the sets $S_2 \cap S_1$, $S_3 \cap S_1$, and $S_4 \cap
  S_1$, and their images $P_2 \cap P_1$, $P_3 \cap P_1$, and $P_4 \cap
  P_1$ respectively.  It is clear that this path labeling preserves
  pairwise intersection cardinalities.  In this case, it is clear that
  the 3 sets are assigned paths in the path $P_1$.  Now by applying
  Lemma~\ref{lem:intersection-cardinality}, it follows that
  $|\cap_{i=1}^4 S_i| =| \cap_{i=1}^4 P_i|$.
\end{proof}
\begin{corollary} 
  \label{corr:4-way} 
  Let $(\cF, \cl)$ be an \ICPPL and $\cl(S_i)=P_i$, $S_i \in \cF$, $1
  \leq i \leq 4$. If $P_i$, $1
  \leq i \leq 4$ are such that $P_1 \cap P_2$, $P_1
  \setminus P_2$, $P_2 \setminus P_1$ are also paths, then:
  \begin{align*}
   |(S_1 \setminus S_2) \cap S_3 \cap S_4| &= |(P_1 \setminus P_2) \cap P_3 \cap P_4|\\
   |(S_2 \setminus S_1) \cap S_3 \cap S_4| &= |(P_2 \setminus P_1) \cap P_3 \cap P_4| 
  \end{align*}
\end{corollary}
\begin{proof}
  It is clear that $|(P_1 \setminus P_2) \cap P_3 \cap P_4| =
  |\displaystyle \cap_{i=1}^4 P_i| - | P_2 \cap P_3 \cap P_4|$.  Since
  we have an ICPPL, from Lemma~\ref{lem:4-way}, it follows that the
  intersection cardinalities are preserved.
\end{proof}
\begin{lemma}
  \label{lem:priv-card}
  Let $(\cF, \cl)$ be an \ICPPL and $\cl(S) = P$ for some $S \in \cF$.  If sets
  $S_{priv}$ and $P_{priv}$ contain only those elements of $S$ and $P$
  that occur in no set other than $S$ and $P$, respectively, then
  $|S_{priv}|=|P_{priv}|$.
\end{lemma}
\begin{proof}
  Consider the set system $ \cF' = \{S_i \cap S \mid S_i \in \cF, i
  \in [n]\}$ and path labeling $\cl'(S_i \cap S) = P_i \cap P$.  Since
  $(\cF, \cl)$ is an \ICPPL, it can be easily verified that $(\cF',
  \cl')$ is an ICPIA on path (interval) $P$.  Let $S_{two}$ and
  $P_{two}$ be the elements of $S$ and $P$ which are in one {\em
    other} set and path, respectively \ie $S_{two} = S \cap (\cup_{S_i
    \ne S} S_i)$ and $P_{two}$ is defined analogously. It may be noted
  that $S_{two} = \supt{\cF'}$ and $P_{two} = \supt{\cF'^{\cl'}}$.
  From Lemma~\ref{lem:icpia} it follows that there is a bijection
  $\phi$ from $S_{two}$ to $P_{two}$ such that for each $i$, the image
  of $S_i \cap S$ under $\phi$ is $P_i \cap P$ (definition of
  feasibility).  Moreover, this means $|S_{two}| = |P_{two}|$. By
  definition, $S_{priv} = S \setminus S_{two}$ and $P_{priv} = P \setminus P_{two}$.
  Thus $|S_{priv}| = |S \setminus S_{two}| = |P \setminus
  P_{two}|=|P_{priv}|$.  
\end{proof}

Now we will prove the effectiveness of Algorithm~\ref{al:filterone}.
Lemma~\ref{lem:feasible} proves that \filteri does not destroy
the feasibility of the input, if it is indeed feasible. Following
this, Lemma~\ref{lem:invar1} and Lemma~\ref{lem:noexit1} show that if
the input is an ICPPL the result of \filteri is also an ICPPL
and they both have the same set of feasibility hypergraph
isomorphisms.

\begin{lemma}
  \label{lem:feasible}
  In Algorithm~\ref{al:filterone}, if input $(\cF, \cl)$ is a feasible path
  labeling then at the end of $j$th iteration of the {\bf while}
  loop, $j \ge 0$, $(\cF_j, \cl_j)$ is a feasible path labeling.
\end{lemma}
\begin{proof}\thesisspacing
  We will prove this by mathematical induction on the number of
  iterations. The base case $(\cF_0, \cl_0)$ is feasible since it is
  the input itself which is given to be feasible. Assume the lemma is
  true till $j-1$th iteration \ie every hypergraph isomorphism $\phi:
  supp\left(\cF_{j-1}\right) \rightarrow V\left(T \right)$ that
  defines $(\cF, \cl)$'s feasibility, is such that the induced path
  labeling on $\cF_{j-1}$, say denoted by $\cp_{\phi[{\cF_{j-1}}]}$,
  is equal to $\cl_{j-1}$.
% Note that any hypergraph isomorphism of
% $\cF$ (say, one that defines the feasibility of $(\cF,\cl)$) can be
% ``projected'' to $\cF_{j-1}$ since their supports are equal. If $\phi$
% is a hypergraph isomorphism of to $(\cF, \cl)$, then 
  We will prove that $\phi$ is also the bijection that makes $(\cF_j,
  \cl_j)$ feasible. Note that $supp(\cF_{j-1}) = supp(\cF_{j})$ since
  the new sets in $\cF_j$ are created from basic set operations to the
  sets in $\cF_{j-1}$ adding or removing no elements. For the same
  reason and $\phi$ being a bijection, it is clear that when applying
  the $\phi$-induced path labeling on $\cF_j$, $
  \cp_{\phi[{\cF_{j}}]}(S_1 \setminus S_2) =
  \cp_{\phi[{\cF_{j-1}}]}(S_1) \setminus
  \cp_{\phi[{\cF_{j-1}}]}(S_2)$. Now observe that $ \cl_j(S_1
  \setminus S_2) = \cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2) =
  \cp_{\phi[{\cF_{j-1}}]}(S_1) \setminus
  \cp_{\phi[{\cF_{j-1}}]}(S_2)$. Thus the induced path labeling
  $\cp_{\phi[{\cF_{j}}]} = \cl_{j}$.  %\qed
\end{proof}

@@

    This invariant is ensured at the end of every iteration due to
    Lemma~\ref{lem:4-way} and Corollary~\ref{corr:4-way}.
   % \begin{enumerate} [{Case 3.}1:]
    % \item {\em $R$, $R'$ and $R''$ are not new sets.} Trivially true
    %   by induction hypothesis.
    % \item {\em Only one, say $R$, is a new set.}  If $R = S_1 \cap
    %   S_2$, from Lemma~\ref{lem:fourpaths} and invariant III
    %   hypothesis, this case is proven. Similarly if $R$ is any of
    %   the other new sets, the case is proven by also using Lemma
    %   ~\ref{lem:setminuscard}.
    % \item {\em At least two of $R, R', R''$ are new sets.}  The new
    %   sets are disjoint hence this case is vacuously true.
    % \end{enumerate}
  \end{enumerate} 
\end{proof}

\begin{lemma}
  \label{lem:noexit1}
  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{al:filterone} is
  feasible, then the set of hypergraph isomorphism functions that
  defines $(\cF, \cl)$'s feasibility is the same as the set that
  defines $(\cF_j, \cl_j)$'s feasibility, if any.%   Secondly, for any
%   iteration $j > 0$ of the {\em \bf while} loop, the {\em \bf exit}
%   statement in line~\ref{ln:exit1} will not execute.
\end{lemma}
\begin{proof}\thesisspacing
  Since $(\cF,\cl)$ is feasible, by Lemma~\ref{lem:feasible}
  $(\cF_j,\cl_j)$ for every iteration $j > 0$ is
  feasible.  % Therefore,
  % every hypergraph isomorphism $\phi: supp(\cF) \rightarrow V(T)$
  % that induces $\cl$ on $\cF$ also induces $\cl_{j-1}$ and $\cl_{j}$
  % on $\cF_{j-1}$ and $\cF_{j}$ respectively, i.e.,
  % $\cl_{\phi[\cF_{j-1}]} = \cl_{j-1}$ and $\cl_{\phi[\cF_j]} =
  % \cl_j$. Thus it can be seen that for all $x \in supp(\cF)$, for
  % all $v \in V(T)$ the following hold true.
  Also, every hypergraph isomorphism $\phi: supp(\cF) \rightarrow
  V(T)$ that induces $\cl$ on $\cF$ also induces $\cl_{j}$ on
  $\cF_{j}$, i.e., $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
  that for all $x \in supp(\cF)$, for all $v \in V(T)$, if $(x,v) \in
  \phi$ then $v \in \cl_{j}(S)$ for all $S \in \cF_{j}$ such that $x
  \in S$.
  % the following hold true.
  % \begin{enumerate}[i. ]
  % \item If $(x,v) \in \phi$ then $v \in \cl_{j-1}(S)$ for all $S \in
  %   \cF_{j-1}$ such that $x \in S$.
  % \item If $(x,v) \in \phi$ then $v \in \cl_{j}(S)$ for all $S \in
  %   \cF_{j}$ such that $x \in S$
  % \end{enumerate}
  In other words, \filteri outputs a filtered path labeling that
  ``preserves''
  hypergraph isomorphisms of the original path labeling.
%   Secondly, line~\ref{ln:exit1} will execute \iff the exit condition in
%   line~\ref{ln:3waycheck}, i.e. failure of three way intersection
%   preservation, becomes true in any iteration of the {\em \bf while}
%   loop.  Due to Lemma~\ref{lem:invar1} Invariant IV, the exit
%   condition does not occur if the input is a feasible ICPPL.%\qed
  % such that $\phi(x) = v$ where $v$ is the leaf considered in the
  % first iterations of while. Clearly, $\phi$ is a renaming of
  % vertices in hypergraph $\cF$ to those in hypergraph
  % $\cF^\cl$. Thus the following facts can be observed in every
  % iteration of the loop.
  % \begin{enumerate}[\hspace{2mm}i. ] \vspace{\topshrink}
  % \item all intersection cardinalities are preserved in this path
  %   labeling \vspace{\topshrink}
  % \item element $x$ is exclusive in a hyperedge in $\cF$ since $v$
  %   is exclusive in a hyperedge in $\cF^\cl$.
  % \end{enumerate}
  % Thus the exit condition is never rendered true after $x$ and $v$
  % are removed from their respective hyperedges. %\qed

@@

As a result of \filteri each leaf $v$ in $T$ is such that there
is exactly one set in $\cF$ with $v$ as a vertex in the path assigned
to it.  \filterii, in Algorithm~\ref{al:filtertwo}, identifies
elements in $supp(\cF)$ whose images are leaves in a hypergraph
isomorphism if one exists.  Let $S \in \cF$ be such that $\cl(S)$ is a
path with leaf and $v \in V(T)$ is the unique leaf incident on it.  We
define a new path labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) =
\set{v}$ where $x$ an arbitrary element from $S \setminus \bigcup_{\hS
  \ne S} \hS$. In other words, $x$ is an element present in no other
set in $\cF$ except $S$. This is intuitive since $v$ is present in no
other path image under $\cl$ other than $\cl(S)$.  The element $x$ and
leaf $v$ are then removed from the set $S$ and path $\cl(S)$
respectively. After doing this for all leaves in $T$, all path images
in the new path labeling $\cl_{new}$ except leaf labels (a path that
has only a leaf is called the {\em leaf label} for the corresponding
single element hyperedge or set) are paths from a new pruned tree $T_0
= T \setminus \{v \mid v \text{ is a leaf in }
T\}$. Algorithm~\ref{al:filtertwo} is now presented with details.

@@

% \begin{lemma}
%   \label{lem:setminuscard}
%   If $\cA$ is an ICPPL, and $(S_1, P_1),(S_2, P_2),(S_3, P_3) \in
%   \cA$, then $|S_1 \cap (S_2 \setminus S_3)| = |P_1 \cap (P_2
%   \setminus P_3)|$.
% \end{lemma}
% \begin{proof}\thesisspacing
%   $|S_1 \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| =
%   |S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to conditions (ii)
%   and (iii) of ICPPL, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1
%   \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3|
%   = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is proven. %\qed
% \end{proof}
